---
title: "CondDensReg"
output: rmarkdown::html_vignette
bibliography: references.bib 
vignette: >
  %\VignetteIndexEntry{CondDensReg}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
The vignette illustrates how to use the different functions included in the **CondDensReg**-package. The package is designed for the regression of conditional densities based on the approach by @maier2024 which treats the conditional densities as elements of the Bayes Hilbert space. The approach is able to handle continuous and discrete densities as well as densities consisting of discrete and continuous components (mixed case).

## Load the data
We use the data set <tt> birthDistribution </tt> from the package **{CondDensReg}**, containing densities of live births in Germany over the months per year (1950-2019) and sex (male and female), resulting in 140 densities.
It is a list with the following elements:
<ul>
* <tt>birth_densities</tt>:
A 140 x 12 matrix containing the birth densities in its rows. The first 70 rows correspond to male newborns, the second 70 rows to female ones. Within both of these, the years are ordered increasingly (1950-2019).
* <tt>birth_densities_clr</tt>:
A 140 x 12 matrix containing the clr transformed densities in its rows. Same structure as <tt>birth_densities</tt>.
* <tt>sex</tt>:
A factor vector of length 140 with levels <tt>"m"</tt> (male) and <tt>"f"</tt> (female), corresponding to the sex of the newborns for the rows of <tt>birth_densities</tt> and <tt>birth_densities_clr</tt>. The first 70 elements are <tt>"m"</tt>, the second 70 <tt>"f"</tt>.
* <tt>year</tt>:
A vector of length 140 containing the integers $1950, \ldots, 2019, 1950, \ldots, 2019$, corresponding to the years for the rows of <tt>birth_densities</tt> and <tt>birth_densities_clr</tt>.
* <tt>month</tt>:
A vector containing the integers from 1 to 12, corresponding to the months for the columns of <tt>birth_densities</tt> and <tt>birth_densities_clr</tt> (domain $\mathcal{T}$ of the (clr-)densities).
</ul>

Instead of the list format, we need a data frame of observations
```{r setup, warning=FALSE, message=FALSE, echo=TRUE}
# load CondDensReg
library(CondDensReg)
# load birth_densities
data("data_age_birth", package = "CondDensReg")
```

## Preprocess the data into count data
```{r, warning=FALSE, message=FALSE, echo=TRUE}
# transform list data in data frame of individual observations

```
## Plot preprocessed data

```{r, warning=FALSE, message=FALSE, echo=TRUE}
library(dplyr)
# plot


```

## Estimate the conditional densities

## Plot different aspects of the estimated model

## References 

